Competing Bandits: \\ The Perils of Exploration under Competition

### 2019-02-14 paper/content/non_greedy_choice.tex
Some edits.

In this section we consider an extension of our base model where the agents' choice model is no longer deterministic. Recall that in our base model agents deterministically choose the firm with the higher reputation score. We define this choice model as \textit{HardMax} (\HM). We study a small deviation from this choice model where the agent now selects between the firms uniformly with probability $0 < \epsilon < 1$ and takes the firm with the higher reputation score with probability $1 - \epsilon$. Following \cite{mansour2018competing} we define this choice model as \textit{HardMax with random} (\HMR).

In Section \ref{sec:competition} we show that in the permanent duopoly
case exploration can lead to a death spiral which eventually starves
the firm of agents. However, giving one firm a small head start or
enough free agents via the warm start incentivizes it to play \TS
since it could recover the reputation costs it incurred from early
exploration. An interpretation of \HMR~ is that instead of
concentrating the free agents as arriving in the beginning of the game
instead they are dispersed throughout the game.

\footnotesize
\begin{table*}[t]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
  \hline
  & \multicolumn{3}{c||}{Heavy-Tail (\HMR)}
  & \multicolumn{3}{c|}{Heavy-Tail (\HM)}\\
  \hline
  & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG 
 & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG  \\
  \hline
$T = 2000$
 & \makecell{ \textbf{0.43} $\pm$ 0.02 \\Var: 0.15 } 
  & \makecell{ \textbf{0.44} $\pm$ 0.02 \\Var: 0.15 } 
  & \makecell{ \textbf{0.6} $\pm$ 0.02 \\Var: 0.1 }
  %%
 &  \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
  & \makecell{ \textbf{0.28} $\pm$ 0.03 \\Var: 0.19 } 
  & \makecell{ \textbf{0.63} $\pm$ 0.03 \\Var: 0.18 }
    \\
\hline
  $T= 5000$ 
   & \makecell{ \textbf{0.66} $\pm$ 0.01 \\Var: 0.056 } 
  & \makecell{ \textbf{0.59} $\pm$ 0.02 \\Var: 0.092 } 
  & \makecell{ \textbf{0.56} $\pm$ 0.02 \\Var: 0.098 } 
  %%
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 } 
 & \makecell{ \textbf{0.62} $\pm$ 0.03 \\Var: 0.19 }
 \\ 
  \hline
  $T = 10000$
  & \makecell{ \textbf{0.76} $\pm$ 0.01 \\Var: 0.026 } 
 & \makecell{ \textbf{0.67} $\pm$ 0.02 \\Var: 0.067 } 
 & \makecell{ \textbf{0.52} $\pm$ 0.02 \\Var: 0.11 }
 %%
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.21 } 
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.2 } 
  & \makecell{ \textbf{0.6} $\pm$ 0.03 \\Var: 0.2 }
  %%
  \\
   \hline
\end{tabular}
\normalsize
\caption{Performance Comparison between \HM and \HMR Choice on the Heavy-Tail MAB instance. Each cell describes the market shares in a game between two algorithms, call them Alg1 vs. Alg2, at a particular value of $t$. Line 1 in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\epsilon = 0.1$ in the \HMR choice model and $T_0 = 20$.}
\label{tab:non_greedy_table}
\end{table*}

@@

\begin{finding}\label{find:non_greedy_choice}
\textit{Under the \HMR~ choice model \TS is weakly dominant as long as $T$ is sufficiently large. \swdelete{However, both $T$ and $\epsilon$ have to be relatively large for this to occur.}}
\end{finding}

Table \ref{tab:non_greedy_table} shows the average market shares under
the \HM~ vs \HMR~ choice model. \swedit{In contrast to the \HM~ model,
  \TS becomes weakly dominant under the \HMR model, as $T$ gets
  sufficiently large.} These results hold across all problem
instances.\footnote{The results here are pulled using different \MRV
  and realizations than the results pulled previously (due to the
  larger $T$). However, they are drawn from the same prior instances
  and so qualitatively are the same but the quantitative results are
  not directly comparable to those from the previous
  section.\swcomment{not sure we need this}} The intuition for why \TS
should become the dominant strategy eventually is simple. The
consistent stream of random agents ensures that each principal is
chosen at least $\Omega(\epsilon t)$ times at every time step $t$. As
a result, each algorithm should eventually converge to its asymptotic
performance in isolation.

However, it takes a significant amount of randomness and a relatively large time horizon for this convergence to occur. Even with $T = 10000$ and $\epsilon = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance. Table \ref{tab:non_greedy_table} also shows that another difference between the two choice models is that \HMR leads to lower variance in market shares across simulations compared to \HM.

>>>>

Let us consider an extension in which the agents' choice is no longer deterministic. Recall that in our main model agents deterministically choose the firm with the higher reputation score; call this choice rule \emph{HardMax} (\HM). Now, we introduce some randomness: each agent selects between the firms uniformly with probability $\eps\in (0,1)$, and takes the firm with the higher reputation score with the remaining probability; call this choice rule \emph{HardMax with randomness} (\HMR).

One can view \HMR as a version of ``warm start", where a firm receives some customers without competition, but these customers are dispersed throughout the game. The expected duration of this ``dispersed warm start" is $\eps T$. If this quantity is large enough, we expect better algorithms to reach their long-term performance and prevail in competition. We confirm this intuition; we also find that this effect is negligible for smaller (but still relevant) values of $\eps$ or $T$.

\OMIT{In Section \ref{sec:competition} we show that in the permanent duopoly
case exploration can lead to a death spiral which eventually starves
the firm of agents. However, giving one firm a small head start or
enough free agents via the warm start incentivizes it to play \TS
since it could recover the reputation costs it incurred from early
exploration. An interpretation of \HMR~ is that instead of
concentrating the free agents as arriving in the beginning of the game
instead they are dispersed throughout the game.}

\footnotesize
\begin{table*}[t]
\centering
\begin{tabular}{|c|c|c|c||c|c|c|}
  \hline
  & \multicolumn{3}{c||}{Heavy-Tail (\HMR)}
  & \multicolumn{3}{c|}{Heavy-Tail (\HM)}\\
  \hline
  & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG
 & \TS vs \DG & \TS vs \DEG  & \DG vs \DEG  \\
  \hline
$T = 2000$
 & \makecell{ \textbf{0.43} $\pm$ 0.02 \\Var: 0.15 }
  & \makecell{ \textbf{0.44} $\pm$ 0.02 \\Var: 0.15 }
  & \makecell{ \textbf{0.6} $\pm$ 0.02 \\Var: 0.1 }
  %%
 &  \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 }
  & \makecell{ \textbf{0.28} $\pm$ 0.03 \\Var: 0.19 }
  & \makecell{ \textbf{0.63} $\pm$ 0.03 \\Var: 0.18 }
    \\
\hline
  $T= 5000$
   & \makecell{ \textbf{0.66} $\pm$ 0.01 \\Var: 0.056 }
  & \makecell{ \textbf{0.59} $\pm$ 0.02 \\Var: 0.092 }
  & \makecell{ \textbf{0.56} $\pm$ 0.02 \\Var: 0.098 }
  %%
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 }
 & \makecell{ \textbf{0.29} $\pm$ 0.03 \\Var: 0.2 }
 & \makecell{ \textbf{0.62} $\pm$ 0.03 \\Var: 0.19 }
 \\
  \hline
  $T = 10000$
  & \makecell{ \textbf{0.76} $\pm$ 0.01 \\Var: 0.026 }
 & \makecell{ \textbf{0.67} $\pm$ 0.02 \\Var: 0.067 }
 & \makecell{ \textbf{0.52} $\pm$ 0.02 \\Var: 0.11 }
 %%
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.21 }
  & \makecell{ \textbf{0.3} $\pm$ 0.03 \\Var: 0.2 }
  & \makecell{ \textbf{0.6} $\pm$ 0.03 \\Var: 0.2 }
  %%
  \\
   \hline
\end{tabular}
\normalsize
\caption{Performance Comparison between \HM and \HMR Choice on the Heavy-Tail MAB instance. Each cell describes the market shares in a game between two algorithms, call them Alg1 vs. Alg2, at a particular value of $t$. Line 1 in the cell is the market share of Alg 1: the average (in bold) and the 95\% confidence band.
%For example, the cell in the top left indicates that TS gets on average 64\% of the market when played against DG.
Line 2 specifies the variance of the market shares across the simulations. The results reported here are with $\eps = 0.1$ in the \HMR choice rule and $T_0 = 20$.}
\label{tab:non_greedy_table}
\end{table*}

@@

\begin{finding}\label{find:non_greedy_choice}
\textit{\TS is weakly dominant under the \HMR~ choice rule, if and only if $\eps T$ is sufficiently large.}
\end{finding}

Table \ref{tab:non_greedy_table} shows the average market shares under
the \HM~ vs \HMR~ choice rule. In contrast to the \HM~ model,
  \TS becomes weakly dominant under the \HMR model, as $T$ gets
  sufficiently large. These findings hold across all problem
instances.\footnote{The results here are pulled using different \MRV
  and realizations than the results pulled previously (due to the
  larger $T$). However, they are drawn from the same prior instances
  and so qualitatively are the same but the quantitative results are
  not directly comparable to those from the previous
  section.\swcomment{not sure we need this}}

\OMIT{The intuition for why \TS
should become the dominant strategy eventually is simple. The
consistent stream of random agents ensures that each principal is
chosen at least $\Omega(\eps t)$ times at every time step $t$. As
a result, each algorithm should eventually converge to its asymptotic
performance in isolation.}

However, it takes a significant amount of randomness and a relatively large time horizon for this effect to take place. Even with $T = 10000$ and $\eps = 0.1$ we see that \DEG still outperforms \DG on the Heavy-Tail MAB instance as well as that \TS only starts to become weakly dominant at $T = 10000$ for the Uniform MAB instance. Table \ref{tab:non_greedy_table} also shows that another difference between the two choice rules is that \HMR leads to lower variance in market shares across simulations compared to \HM.
